Algorithm explanation

### 2013-08-28 article/conclusion.tex
Spell check

To accomplish this goal on Craig interpolation over linear
arithmetics, we first explained a method to reduce interpolation into
linear constraint generation by using Farkas's Lemma.  This method
preserves a set of solutions during the computation process so to make
it possible to choose a small solution later. This enabled a Horn
clause solver to build small solutions when it uses the method as a
subroutine due to the ability to choose a common predicate among
different interpolating problems. Additionally, we mentioned the
possibility of chosing small solutions in interpolating general
formulas by wisely choose the order to combine set of solutions during
the computation. In the end we extended the method to solve symmetric
interpolation problems to perform predicate discovery at multiple
locations in the program at the same time.

>>>>

To accomplish this goal on Craig interpolation over linear
arithmetics, we first explained a method to reduce interpolation into
linear constraint generation by using Farkas's Lemma.  This method
preserves a set of solutions during the computation process so to make
it possible to choose a small solution later. This enabled a Horn
clause solver to build small solutions when it uses the method as a
subroutine due to the ability to choose a common predicate among
different interpolating problems. Additionally, we mentioned the
possibility of choosing small solutions in interpolating general
formulas by wisely choose the order to combine set of solutions during
the computation. In the end we extended the method to solve symmetric
interpolation problems to perform predicate discovery at multiple
locations in the program at the same time.

### 2013-08-28 article/experiment.tex
Spell check

There are some interpolating theorem provers known.
\begin{itemize}
\item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
  arithmetic reasonings and SAT-based logical reasonings.
\item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
  proof-based method.
\item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
 from constraints by linear arithmetic reasonings. It only supports
 conjunctive problems.
\end{itemize}

In our experiment, we compared the performance of our algorithm with
CSIsat, that is mainly used as a subprocedure of MoCHi.  Note that
CSIsat may return different interpolants for an interpolating problem
because of the randomized algorithm.

>>>>

There are some interpolating theorem provers known.
\begin{itemize}
\item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
  arithmetic reasoning and SAT-based logical reasoning.
\item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
  proof-based method.
\item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
 from constraints by linear arithmetic reasoning. It only supports
 conjunctive problems.
\end{itemize}

In our experiment, we compared the performance of our algorithm with
CSIsat, that is mainly used as a sub procedure of MoCHi.  Note that
CSIsat may return different interpolants for an interpolating problem
because of the randomized algorithm.

### 2013-08-28 article/future.tex
Formatting for final
Spell check

Our proposed algorithms have still numbers of rooms for further
improvement.  Based on the experiment result from the previous
chapter, here we give raise some analaysis and discussion for them.

@@

In the preliminary expriment before incorporating into MoCHi, although
we have tried the quantifier elimination over integer theory, the
result was unsatisfactory in the sense that simplified formulas had
disjunctions.  It caused the loss of efficiency in solving constraints
and generating unsatisfiable cores.

We have also tried the quantifier elimination over real space, and the
result was better than the integer version.  However, because our
algorithm first treats all linear inequalities over integer space,
more study on the gap betweem these two are required when we consider
the completeness.

@@

\paragraph{Generation and choice of unsatisfiable core}
We can efficiently build linear constraints for the given $G$ after
the transformation in $\transGraph$~procedure by appropriate inference
from the previous constraint and the SMT query result.  Our Horn
clause solver make use of Z3 to obtain an unsatisfiable core from the
linear constraints.  However the extraction of an unsatisfiable core
from the proof sums up to a large portion of the whole computation
time in our algorithm.

Although our implementation in this work did not consider such the way
of optimization to reduce the cost of linear constraints queries, it
is necessary to reduce the number of them for reduce the computation
cost.  For instance, the optimization is possible by appropriatealy
updating $V_\star$ in $\solveHorn$~procedure since the vertices that
are already examined for split does not need to be considered again.

The choice of subset of constraints for an unsatisfiable core is also
important.  In some cases, there exists different multiple cores and
the algorithm returns a different shape of the solution in relation to
the choice of them.  Thus, we need further study to optimize the order
of choice of multiple unsatisfiable cores if ones exists.

\paragraph{Separation of edges}
To transform the graph by $\transGraph$~procedure, there may be
multiple transformations possible.  For example, consider a predicate
variable $P$ and edges $\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$
that go out from the vertex.  If an unsatisfiable core contains a set
of edges $\left\lbrace e_1, e_2 \right\rbrace$,
$\transGraph$~procedure determines to split those edges by duplicating
$P$.  However at this point, $e_3$ and $e_4$ are free to be connected
to any duplication of $P$.  Alghough our current algorithm split
$\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$ into four duplications
$\left\lbrace e_1 \right\rbrace$ to $\left\lbrace e_4 \right\rbrace$,
it is possible to split into $\left\lbrace e_1, e_3\right\rbrace$ and
$\left\lbrace e_2, e_4 \right\rbrace$ for example.  By intellilgently
choose this, we may be able to obtain much simpler solution than the
current algorithm.

We suggest to reduce this edge split problem into a graph coloring
problem.  By generating a graph which a set of vertices have an edge
of Horn graph as a label and solve a coloring problem on it, the
algorithm can wisely determine the minuminm number of duplications.

@@

\paragraph{Applications for other work}
In our research, we put a special focus on solving Horn clauses for
dependent type inference in higher-order program verification.
However, solving Horn clauses are becoming common to describe various
kinds of programs' behavior. For instance in verifying multi-threaded
programs, Horn clauses can describe the conditions of data exchange
between threads, and confirm the validity of semaphores for mutural
exclusion.  We expect our algorithm achieve better performance in
other work when customized into different contexts.

By puting such effort for further improvement on our algorithm, we can
conclude that the algorithm may be able to achive better performance
in Horn clause solving, and become a necessary tool for program
verification.

>>>>

Our proposed algorithms have still numbers of rooms for further
improvement.  Based on the experiment result from the previous
chapter, here we give raise some analysis and discussion for them.

@@

In the preliminary experiment before incorporating into MoCHi, although
we have tried the quantifier elimination over integer theory, the
result was unsatisfactory in the sense that simplified formulas had
disjunctions.  It caused the loss of efficiency in solving constraints
and generating unsatisfiable cores.

We have also tried the quantifier elimination over real space, and the
result was better than the integer version.  However, because our
algorithm first treats all linear inequalities over integer space,
more study on the gap between these two are required when we consider
the completeness.

@@

\paragraph{Generation and choice of unsatisfiable core}
We can efficiently build linear constraints for the given $G$ after
the transformation in $\transGraph$~procedure by appropriate inference
from the previous constraint and the SMT query result.  Our Horn
clause solver makes use of Z3 to obtain an unsatisfiable core from the
linear constraints.  However the extraction of an unsatisfiable core
from the proof sums up to a large portion of the whole computation
time in our algorithm.

Although our implementation in this work did not consider such the way
of optimization to reduce the cost of linear constraints queries, it
is necessary to reduce the number of them for reduce the computation
cost.  For instance, the optimization is possible by appropriately
updating $V_\star$ in $\solveHorn$~procedure since the vertices that
are already examined for split does not need to be considered again.

The choice of subset of constraints for an unsatisfiable core is also
important.  In some cases, there exists different multiple cores and
the algorithm returns a different shape of the solution in relation to
the choice of them.  Thus, we need further study to optimize the order
of choice of multiple unsatisfiable cores if ones exist.

\paragraph{Separation of edges}
To transform the graph by $\transGraph$~procedure, there may be
multiple transformations possible.  For example, consider a predicate
variable $P$ and edges $\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$
that go out from the vertex.  If an unsatisfiable core contains a set
of edges $\left\lbrace e_1, e_2 \right\rbrace$,
$\transGraph$~procedure determines to split those edges by duplicating
$P$.  However at this point, $e_3$ and $e_4$ are free to be connected
to any duplication of $P$.  Although our current algorithm split
$\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$ into four duplications
$\left\lbrace e_1 \right\rbrace$ to $\left\lbrace e_4 \right\rbrace$,
it is possible to split into $\left\lbrace e_1, e_3\right\rbrace$ and
$\left\lbrace e_2, e_4 \right\rbrace$ for example.  By intelligently
choose this, we may be able to obtain much simpler solution than the
current algorithm.

We suggest to reduce this edge split problem into a graph coloring
problem.  By generating a graph which a set of vertices have an edge
of Horn graph as a label and solve a coloring problem on it, the
algorithm can wisely determine the minimum number of duplications.

@@

\paragraph{Applications for other work}
In our research, we put a special focus on solving Horn clauses for
dependent type inference in higher-order program verification.
However, solving Horn clauses are becoming common to describe various
kinds of programs' behavior.  For instance in verifying multi-threaded
programs, Horn clauses can describe the conditions of data exchange
between threads, and confirm the validity of semaphores for mutual
exclusion.  We expect our algorithm achieve better performance in
other work when customized into different contexts.

By putting such effort for further improvement on our algorithm, we can
conclude that the algorithm may be able to achieve better performance
in Horn clause solving, and become a necessary tool for program
verification.

### 2013-08-28 article/horn.tex
Spell check

\begin{align*}
\text{Constraint:} \\
& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
& c_1 + c_{3,2} = 0 \wedge c_2 + c_{4,2} > 0
\\
\text{Unsatisfiable core:} \\
0 \geq c_2 \\
& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \\
& - \lambda_4 = c_{3,2} \wedge \\
& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
& c_1 + c_{3,2} = 0
\end{align*}

@@

@@

After this, we obtain the following constraint:

\begin{align*}
& \lambda_1 - \lambda_2 = c_{1,1} \wedge 0 \geq c_{2,1} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
& \lambda_1 - \lambda_2 = c_{1,2} \wedge 0 \geq c_{2,2} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
& c_{1,1} + c_{1,2} + c_{3,1} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,1} > 0 \wedge \\
& c_{1,1} + c_{1,2} + c_{3,2} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,2} > 0
\end{align*}

By continuing this process, the process in the end generates a
predicate template that the solution will become.  Eventually, we
obtain the following solution.

\begin{align*}
A(x) \text{ : } & x \leq 0 \wedge -x \leq 0 \\
B(x) \text{ : } & x+1 \leq 0 \vee -x+1 \leq 0
\end{align*}

@@

A variable mapping label $\theta$ on edges is used to map the index of
one predicate variable to another.

>>>>

\begin{align*}
\text{Constraint:} \\
TODO \\
\\
\text{Unsatisfiable core:} \\
TODO
\end{align*}

@@

TODO

@@

Finally, we obtain the following constraint:

TODO

The process also generates a predicate template for every predicate
variable that the solution will become.

@@

A variable mapping label $\theta$ on edges is used to map the index of
one predicate variable to another. (TODO?)

(Example?)

### 2013-08-28 article/interpolation.tex
Spell check

In computing the interpolant, we make a linear constraint of a
interpolant by applying Farkas's lemma.  First, we assing weight
parameters to every expression in conjunction form.

@@

Our algorithm aims to construct a relatively small interpolant in a
reasonable computation time for any input.  It also aims to obtain a
common solution among multiple interpolating problems.  When the
algorithm is used as a sub-routine for solving Horn clause satisfying
problems that are made while discovering predicates for abstraction
during the program verification.  These advantages of our algorithm
make it possible to return a relatively small predicates for those
problems. To accomplish these aims, the algorithm preserves a set of
interpolants during the computation, and define and execute operations
over interpolant sets.

@@

Then, a linear combination of $A$'s formula with a weight $\lambda_1,
\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
Note that any assignments to $\lambda_i$ that satisfy the expressions
above will let an expression $I$ to be an interpolant.  Considering
this, it is possible to obtain different interpolants for a
interpolating problem by computing different assignments to
$\lambda_i$.

>>>>

In computing the interpolant, we make a linear constraint of a
interpolant by applying Farkas's lemma.  First, we assign weight
parameters to every expression in conjunction form.

@@

Our algorithm aims to construct a relatively small interpolant in a
reasonable computation time for any input.  It also aims to obtain a
common solution among multiple interpolating problems.  These goals
are set because the algorithm is designed to be used as a sub-routine
for solving Horn clause satisfying problems
that are made while discovering predicates for abstraction
during the program verification.  These advantages of our algorithm
make it possible to return a relatively small predicates for those
problems. To accomplish these aims, the algorithm preserves a set of
interpolants during the computation, and executes operations
over interpolant sets.

@@

Then, a linear combination of $A$'s formula with a weight $\lambda_1,
\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
Note that any assignments to $\lambda_i$ that satisfy the expressions
above will let an expression $I$ to be an interpolant.  Considering
this, it is possible to obtain different interpolants for an
interpolating problem by computing different assignments to
$\lambda_i$.

### 2013-08-28 article/related.tex
Other chapters also separated
Spell check

@@

\paragraph{Development of interpolating theorem provers}
An interpolating prover \textsc{FOCI} \cite{website/foci} was
developed under the comprehensive wrok on quantifier-free interpolant
generation over the theory of linear arithmetic and uninterpreted
function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At the same
time, various applications of interpolants for model checking was
proposed \cite{conf/tacas/McMillan05}. They incorporated the algorithm
into Z3 SMT solver \cite{conf/tacas/MouraB08} and it is possible to
generate an interpolant from proofs generated by
Z3 \cite{conf/fmcad/McMillan11}.

@@

One of them is work done by \cite{conf/pldi/BeyerHMR07} to eliminate
spurious paths that pass over loops.  The method converts a
counterexample from the model checker into a path that is a subgraph
of program's control flow graph, and performs a invariant synthesis
over it.  Here, the control flow graph may contain loops, it is not
suffcieint by the convential interpolation, and rather it required a
disjunctive reasoning.

@@

As another direction of extension, some works are done for solving
Horn clauses with qunatified consequences \cite{conf/sas/BjornerMR13,
conf/cav/BeyenePR13}.  They are used to abstract more complex systems
efficiently.

>>>>

\chapter{Related work}
\label{chap:related}

@@

\paragraph{Development of interpolating theorem provers}
An interpolating prover \textsc{FOCI} \cite{website/foci} was
developed under the comprehensive work on quantifier-free interpolant
generation over the theory of linear arithmetic and uninterpreted
function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At the same
time, various applications of interpolants for model checking was
proposed \cite{conf/tacas/McMillan05}. They incorporated the algorithm
into Z3 SMT solver \cite{conf/tacas/MouraB08} and it is possible to
generate an interpolant from proofs generated by
Z3 \cite{conf/fmcad/McMillan11}.

@@

One of them is work done by \cite{conf/pldi/BeyerHMR07} to eliminate
spurious paths that pass over loops.  The method converts a
counterexample from the model checker into a path that is a subgraph
of program's control flow graph, and performs a invariant synthesis
over it.  Here, the control flow graph may contain loops, it is not
sufficient by the conventional interpolation, and rather it required a
disjunctive reasoning.

@@

As another direction of extension, some works are done for solving
Horn clauses with quantified consequences \cite{conf/sas/BjornerMR13,
conf/cav/BeyenePR13}.  They are used to abstract more complex systems
efficiently.
